%!TEX root = ts.tex

\rSec0[scope]{Scope}

%\pnum
This document describes requirements for implementations of an interface that computer programs written in the C++ programming language could use to invoke algorithms with concurrent execution. The algorithms described by this document are realizable across a broad class of computer architectures.

%\pnum
This document provides important context and specification for
this document. This document is written as a set of changes against that specification.  Instructions to modify or add paragraphs are written as
explicit instructions.

%\pnum
Some of the functionality described by this document might be considered for standardization in a future version of C++, but it is not currently part of ISO/IEC 14882:2020. Some of the functionality in this document might never be standardized, and other functionality might be standardized in a substantially changed form.

%\pnum
The goal of this document is to build widespread existing practice for concurrency in the ISO/IEC 14882:2020 algorithms library. It gives advice on extensions to those vendors who wish to provide them.

\rSec0[refs]{Normative references}

%\pnum
The following documents are referred to in the text in such a way that some or all of their content constitutes requirements of this document. For dated references, only the edition cited applies. For undated references, the latest edition of the referenced document (including any amendments) applies.

\begin{itemize}
\item ISO/IEC 14882:2020, \doccite{Programming Languages --- C++}
\end{itemize}

%\pnum
%Unless otherwise specified, the whole of the ISO/IEC 14882:2020 library introduction (\CppXref{16}) is included into this Technical Specification by reference.


\rSec0[defs]{Terms and definitions}
%\pnum
\indextext{definitions|(}%
No terms and definitions are listed in this document. ISO and IEC maintain
terminological databases for use in standardization at the following addresses:

\begin{itemize}
    \item IEC Electropedia: available at https://www.electropedia.org/
    \item ISO Online browsing platform: available at https://www.iso.org/obp
\end{itemize}

\rSec0[general]{General}

\rSec1[general.compliance]{Implementation compliance}
%\pnum
Conformance requirements for this document are those defined in \CppXref{4.1}, as applied to a merged document consisting of C++20 amended by this document.
%%\begin{note}

NOTE Conformance is defined in terms of the behaviour of programs.%%\end{note}

\rSec1[general.namespaces]{Namespaces and headers and modifications to standard classes}
%\pnum
Since the extensions described in this document are experimental and not part of the ISO/IEC 14882:2020 library, they are not declared directly within namespace \tcode{std}. Unless otherwise specified, all components described in this document either:

\begin{itemize}
    \item modify an existing interface in the ISO/IEC 14882:2020 library in-place,
    \item are declared in a namespace whose name appends \tcode{::experimental::concurrency\_v2} to a namespace defined in the ISO/IEC 14882:2020 library, such as \tcode{std}, or
    \item are declared in a subnamespace of a namespace described in the previous bullet, whose name is not the same as an existing subnamespace of namespace \tcode{std}.
\end{itemize}

%\pnum
Whenever an unqualified name is used
in the specification of a declaration \tcode{D},
its meaning is established
in accordance with (\CppXref{4.1.2}) by performing unqualified name lookup
in the context of \tcode{D}.
\begin{note}
Argument-dependent lookup is not performed.\end{note}
Similarly, the meaning of a \grammarterm{qualified-id} is established
in accordance with performing qualified name lookup
in the context of \tcode{D}.
\begin{note}
Operators in expressions are not so constrained.\end{note}

%\pnum
Table~\ref{tab:info.headers} shows the headers described in this document
%%shall import the contents of %%\tcode{::std::experimental::concurrency::v2} into
%%\tcode{::std::experimental::concurrency_v1} as if by:

%%\begin{codeblock}
%%namespace std::experimental::concurrency_v1 {
%%  inline namespace v2 {}
%%}
%%\end{codeblock}


\begin{floattable}{\textbf{C++ library headers}}{tab:info.headers}
{l}
\topline
\tcode{<experimental/asymmetric_fence>} \\
\tcode{<experimental/bytewise_atomic_memcpy>} \\
\tcode{<experimental/hazard_pointer>} \\ 
\tcode{<experimental/rcu>} \\
\tcode{<experimental/synchronized_value>} \\

\end{floattable}

\rSec1[general.feature.test]{\textbf{Feature-testing recommendations}}
%\pnum
An implementation that provides support for this document should define each feature test macro defined in \tref{intro.features1} and \tref{intro.features2} if no associated headers are indicated for that macro, and if associated headers are indicated for a macro, that macro is defined after inclusion of one of the corresponding headers specified in \tref{intro.features1} and \tref{intro.features2}.

%%\begin{floattable}{\textbf{Feature-test macros name}}{tab:intro.features}
%%{lllll}
%%\topline
%%\lhdr {Title} & \lhdr {Subclause} & \chdr{Macro name} & %%\rhdr{Value} & \rhdr{Header} \\
%%{ll}
%%\lhdr {Title} &
%%\rhdr {Subclause \\ Macro name \\ Value \\ Header} \\

%%\capsep
%%\tcode{__cpp_concurrency\_v2}  & \tcode{\tsver}   &  none \\
%%\tcode{Hazard pointers} & \tcode{5.2} & %%\tcode{__cpp_lib_experimental_hazard_pointer}  & \tcode{\tsver}  %%& \tcode{<experimental/hazard_pointer>} \\
%%\tcode{Read-copy update(RCU)} & \tcode{5.3} & %%\tcode{__cpp_lib_experimental_rcu}  & \tcode{\tsver}  & %%\tcode{<experimental/rcu>} \\
%%\tcode{bytewise atomic memcpy} & \tcode{6} & %%\tcode{__cpp_lib_experimental_bytewise_atomic_memcpy}  & %%\tcode{\tsver}  & \tcode{<experimental/bytewise_atomic_memcpy>} \\
%%\tcode{Asymmetric Fence} & \tcode{7} & %%\tcode{__cpp_lib_experimental_asymmetric_fence}  & \tcode{\tsver} %% & \tcode{<experimental/asymmetric_fence>} \\
%%\tcode{Synchronized Value} & \tcode{8} & %%\tcode{__cpp_lib_experimental_synchronized_value}  & %%\tcode{\tsver}  & \tcode{<experimental/synchronized_value>} \\
%%\end{floattable}

\begin{floattable}{\textbf{Feature-test macros name}}{tab:intro.features1}
{lll}
\topline
\lhdr {Title} & \chdr {Subclause} & \rhdr{Macro name}  \\
%%{ll}
%%\lhdr {Title} &
%%\rhdr {Subclause \\ Macro name \\ Value \\ Header} \\

\capsep
%%\tcode{__cpp_concurrency\_v2}  & \tcode{\tsver}   &  none \\
\tcode{Synchronized Value} & \tcode{5} & \tcode{__cpp_lib_experimental_synchronized_value}   \\
\tcode{Hazard pointers} & \tcode{6.2} & \tcode{__cpp_lib_experimental_hazard_pointer}   \\
\tcode{Read-copy update(RCU)} & \tcode{6.3} & \tcode{__cpp_lib_experimental_rcu}   \\
\tcode{Bytewise atomic memcpy} & \tcode{7} & \tcode{__cpp_lib_experimental_bytewise_atomic_memcpy}   \\
\tcode{Asymmetric Fence} & \tcode{8,33} & \tcode{__cpp_lib_experimental_asymmetric_fence}   \\

\end{floattable}

\begin{floattable}{\textbf{Feature-test macros header}}{tab:intro.features2}
{lll}
\topline
\lhdr {Title} &  \chdr{Value} & \rhdr{Header} \\
%%{ll}
%%\lhdr {Title} &
%%\rhdr {Subclause \\ Macro name \\ Value \\ Header} \\

\capsep
%%\tcode{__cpp_concurrency\_v2}  & \tcode{\tsver}   &  none \\
\tcode{Synchronized Value} &  \tcode{\tsver}  & \tcode{<experimental/synchronized_value>} \\
\tcode{Hazard pointers} &  \tcode{\tsver}  & \tcode{<experimental/hazard_pointer>} \\
\tcode{Read-copy update(RCU)} & \tcode{\tsver}  & \tcode{<experimental/rcu>} \\
\tcode{Bytewise atomic memcpy} &  \tcode{\tsver}  & \tcode{<experimental/bytewise_atomic_memcpy>} \\
\tcode{Asymmetric Fence} &  \tcode{\tsver}  & \tcode{<experimental/asymmetric_fence>} \\

\end{floattable}

% \rSec1[general.plans]{Future plans (Informative)}
% %\pnum
% This clause describes tentative plans for future versions of this document and plans for moving content into
% future versions of ISO/IEC 14882:2020.

% \pnum
%  The C++ committee intends to release a new version of this document approximately every few years, containing
% the concurrency extensions we hope to add to a near-future version of ISO/IEC 14882:2020. Future versions will define their
% contents in \tcode{std::experimental::concurrency\_v3}, \tcode{std::experimental::concurrency\_v4}, etc., with the most recent
% implemented version inlined into \tcode{std::ex\-perimental}.

% \pnum
% When an extension defined in this or a future version of this document represents enough existing practice, it
% will be moved into the next version of ISO/IEC 14882:2020 by removing the \tcode{experimental::con\-currency\_v$N$} segment of its
% namespace and by removing the \tcode{experimental/} prefix from its header's path.

% \rSec1[general.ack]{Acknowledgments}

% This work is the result of a collaboration of researchers in industry and academia. We wish to thank the
% original authors of this document, Michael Wong, Paul McKenney, and Maged Michael, and the editing review team of Jonathan Wakely, Daniel Krügler, and Bryan St. Amour. We also wish to thank people
% who made valuable contributions within and outside these groups, including Jens Maurer, and many others not named
% here who contributed to the discussion.




